\chapter{Introduction}
This chapter presents background, purpose and objectives to make clear the goal of this thesis.


\section{Background}
RFID technology is happening to be very popular these days for various applications such as industrial automation, access control, animal identification, public transport, event ticketing, parking, electronic wallet, goods identification and many more.
The question is how secure this technology is. The answer is, that there are various manufacturers providing RFID (Radio Frequency Identification) devices with various security. Information delivered by the manufacturers about the security should be clear.


\section{Why electronic door lock?}
However, there are very secure door locks, commonly used mechanical door lock has a lot of disadvantages. It's easy to clone keys and it's even possible to open the door without the key. If a key is lost, changing the lock is needed, which could be even impossible in some cases. Somebody who finds the lost key would be able to open the door.
Electronic door lock  might came out with solutions of these problems, however every lock is possible to hack somehow. 
A key of electronic door lock is usually RFID tag or card, but it can be also mobile phone communicating via bluetooth interface etc.
The user's card may be also used in other applications, like electronic purse.
In case that user loses his card, the card can be deleted from the system.
There is also so many advantages like user info can be stored on the RFID card and part of the system might be a server where scanned cards are monitored. 


\section{Objectives}
The purpose of this thesis is to design and assembly an RFID based access control system composed of inexpensive RFID devices used today and examine its security and possibility of hacking it. 
In the conclusions, it's assessed if the information given by the manufacturer gives a clear picture of the security of
his RFID technology.















